PEACE: {\underline P}arallel {\underline E}nvironment for {\underline A}ssembly and {\underline C}lustering of Gene {\underline E}xpression

### 2010-05-11 papers/NAR_web_server_2009/paper.sup.tex
Manuscript update: final(?) version

\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}
\newcommand{\metasim} {{\small MetaSim}}

@@

In {\it Hazelhurst et al.} \cite{Hazelhurst08a} the authors conduct
an investigation of \textsc{wcd} \mc{(version 0.5.1)} against a number of clustering
tools.  As they make a convincing argument that \textsc{wcd} is returning
better results then the tools against which they compare, we limit our
analysis to a comparison of \textsc{peace} against \textsc{wcd} and the \textsc{cap3} assembly
tool \cite{Huang99} -- a tool which implicitly clusters while
performing assembly.  We do so by applying all three tools to a number
of data sets, both simulated and real.

@@

For simulated data sets we use the {\bf ESTsim} tool to generate
simulated EST data sets \cite{Hazelhurst03}, using the collection of
zebra fish genes that served as a basis for the \textsc{wcd} simulations
\cite{Hazelhurst08a}.  In generating the ESTs, ESTsim models three
types of error (general base read errors, errors due to
polymerase decay, and primer interference), and allows those errors to
take the form of substitutions, deletions, and insertions (of bases
and Ns).  See the paper for a discussion of the probability
distributions and default parameters.  For generating our simulated
data, we use the default values for all parameters that are not
explicitly being subjected to variation in our experiments --
paralleling the testing of the \textsc{wcd} tool.

@@

Each estimate given in the main paper, or in the following, is
averaged over 30 trials.  Each trial consists of the application of
all three tools to a simulated data set, the set having been derived
from the application of the ESTsim tool to a set of 100 zebra fish
gene sequences \cite{Hazelhurst03}.  All confidence intervals are
calculated at a 95\% level of significance.

@@

\mc{Our primary measurement tools if measurement are sensitivity, the
Jaccard Index, Type 1 error and Type 2 error
(Figure~3).  Sensitivity is the ratio for true
positives to actual positives ($tp / (tp + fn)$) -- the} \mc{fraction} \mc{of}
\mc{sequences} \mc{from} \mc{the} \mc{same} \mc{transcript} \mc{that} \mc{are} \mc{identified} \mc{as} \mc{such.}  \mc{The}
\mc{{\it Jaccard}} \mc{index} \mc{combines} \mc{both} \mc{sensitivity} \mc{and} \mc{specificity,}
\mc{calculated} \mc{as} \mc{the} \mc{ratio} \mc{of} \mc{true} \mc{positives} \mc{to} \mc{the sum of true
positives, true negatives, and false positives ($tp / (tp+fn+fp)$)
\cite{Hazelhurst08a}.}

@@

\mc{Runtime of the tool is analyzed in Figure~4 of the main paper.  In
Figure~\ref{memory} we present a plot of the memory footprint of each tool
as a function of input size.  Data points were generated by the application of
each tool to simulated sets of a specified size (generated by ESTSim),
measuring the maximum amount of memory required for the run.  We see that while
all tools require memory linear in the input size, \wcd\/ requires
slightly less than does \peace, while both require significantly less
memory than \capthree.  In general we find the footprint of \peace\/
to be relativily small, as compared to other clustering tools
(discussed in {\it Hazelhurst et al.} \cite{Hazelhurst08a}).}

\mc{All runs were conducted on 3.0 GHz Intel Xeon EM64T CPUs with 2 MB
caches and a 800 MHz front side bus, model number Xeon LV 3.0 (2005)}

\begin{figure}[tbp]
\centerline{
\includegraphics[scale=0.35]{pics.d/memory.pdf}
}
\caption{Plots of memory footprint against sequence set size for a
  single processor run on simulated data for each tool: \peace\/
  (blue), \wcd\/ (green) and \capthree\/ (balck), along with the
  lesat-squares fit for each.  For each fit we have an associated
  regression coefficient of $r > 0.99$.}\label{memory}
\end{figure}

\subsection{Raising the Threshold}
\label{sim_thresh}
\mc{ 
  The \wcd\/ tool uses a $d^2$ threshold of $40$ to determine whether
  two sequences are in the same cluster.  As this is equivalent to our
  $T$ threshold for dismissing MST edges, we have set it to the same
  default value when used for Sanger} \mc{sequences.}  \mc{However,} \mc{we} \mc{did}
  \mc{experiment} \mc{with} \mc{higher} \mc{values,} \mc{with} \mc{interesting} \mc{results.  An
  analysis of the preformance of $d^2$ indicated that a $T$ value as high as 130 should not have a
  significant impact on its effectiveness as a discriminator.  And in
  fact this was born out in simulation, as shown in
  Figure~\ref{SeJiT1T2130}.  In contrast with the equivalent results
  for a threshold of $T=40$, as shown in the manuscript, we see
  considerable improvement in \peace\/ over \wcd\/ in all metrics
  except Type 2 error, in which \wcd\/ slightly
  better. Unsurprisingly, this also makes a significant difference in
  the tool's ability to distinguish duplications: we see fairly poor
  performance in Figure~\ref{dups130}.  
}

@@

\begin{figure}
  \centerline{\includegraphics[width=3.35in]{pics.d/SeJiT1T2x130.pdf}}
  \caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
    Type 1 error using the elevated threhold ($T=130$), providing a
    contrast against Figure 3 ($T=40$).  Based on the average over 30
    simulated Sanger Sequence ESTs sets derived from 100 zebra fish
    genes.  Blue/Solid = \peace, Green/Dash = \wcd, Black/Dot-Dash =
    \capthree; vertical tics = 95\% confidence intervals on estimates.
    Intervals are not presented for Type 1 error due to the effective
    lack of variance.}\label{SeJiT1T2130}
\end{figure}

\begin{figure}[tbp]
\centerline{
\includegraphics[scale=0.35]{pics.d/duplicates_130.pdf}
}
\caption{Ability to distinguish duplicates as a function of
  divergence.  Estimates averaged over 30 trials; one trial consists
  of taking a random zebra fish gene, copying it and stochastically changing bases
  at the specified rate, then using the two genes as the bases for
  generating a simulated set. Blue/Solid = \textsc{peace}, Green/Dashed = \textsc{wcd}, Black/Dot-Dashed = \textsc{cap3};
  variance was too small for visible plotting of
  confidence intervals.}\label{dups130}
\end{figure}

\section{Real Data}

@@

\begin{table}[tbp]
\begin{center}
\begin{tabular} {|c|| c | c || c | c |}
\hline
 Benchmark & \begin{tabular}{c} Number of \\ Sequences \end{tabular}
 & \begin{tabular}{c} Number of \\ Bases \end{tabular} & Peace Runtme (s)&
 WCD Runtime (s) \\
\hline
\hline
 Human Cluster & 17,733 & 11.7 Mb & 293 & 804 \\
\hline
 Cotton & 29,995 & 17.1 Mb & 222 & 339 \\
\hline
 Ricinus & 57,690 & 40.9 Mb & 1344 & 2356 \\
\hline
 A076941 & 76,941 & 32.8 Mb & 1166 & 966 \\
\hline
\hline
 A686903 & 686,904 & 294.7 Mb & 9449 & 2481 \\
\hline
\end{tabular}
\end{center}
\caption{Runtime for \peace\/ and \wcd\/ on benchmarks from the
  WCD and EasyCluster studies \cite{Hazelhurst08a, Picardi09}.
  Runtimes for the first four sets are sequential, while that last was
  run over thirty processors.}\label{RunTime}
\end{table}

@@

\begin{table}[tbp]
\begin{center}
\begin{tiny}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime
      (s) \end{tabular} \\
\hline
\hline EasyCluster Human & PEACE & 0.999 & 0.106 & 0.027 & 0.030 & 67 & 2 & 295 \\
\cline{2-9} Benchmark & WCD & 0.998 & 0.672 & 0.144 & 0.044 & 113 & 16 & 804 \\
\cline{2-9} (111 Genes) & Cap3 & 0.657 & 0.643 & 1.000 & 0.001 & 2269 & 1827 & \mbox{NA} \\
\hline
\hline WCD A076941 & PEACE & 0.937 & 0.345 & 0.337 & 0.0342 & 18256 & 8135 & 1293 \\
\cline{2-9} Benchmark & WCD & 0.933 & 0.476 & 0.350 &  0.027 & 18787 & 8553 & 966 \\
\cline{2-9} (13240 genes) & Cap3 & 0.826 & 0.802 & 0.486 & 0.014 & 25042 & 14916 &
\mbox{NA} \\
\hline
\end{tabular}
\end{tiny}
\end{center}
\caption{Comparisons of runs on the EasyCluster human Benchmark
  Dataset and the WCD A076941 Arabidopsis thaliana dataset using the
  standard quality measurements with $T=130$.}\label{SangerReal130}

@@

\begin{table}[htbp]
\begin{center}
\begin{tiny}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime (s) \end{tabular}\\
\hline
\hline   454 & PEACE & 0.871 & 0.809 & 0.186 & 0.049 & 129 & 4 & 129 \\
\cline{2-9}    Simulation & WCD(1) & 0.459 & 0.448 & 0.946 & 0.003 & 599 & 235 & 76 \\
\cline{2-9}  & TGICL & 0.297 & 0.288 & 0.984 & 0.005 & 886 & 0 & 68 \\
\hline
\hline Illumina         & PEACE & 0.384 & 0.343 & 0.230 & 0.018 & 673  & 299 & 3463 \\
\cline{2-9} Simulation  & TGICL & 0.131 & 0.127 & 0.977 & 0.008 & 1522 & 0   & 373 \\
\hline
\end{tabular}
\end{tiny}
\end{center}
\caption{Comparisons of runs on on short read data, as generated by
  \metasim \cite{Richter2008}; each figure is an average over 30 runs.
  For \wcd we use parameters suggested by the authors  (``-H 1 -T 160''). \wcd\/ was
  unable to return any reasonable results for Illumina data, while
  \capthree\/ was unable to return results for either sumulation.  Each
  454 simulation produced 25,000 reads from the 100 zebra-fish
  transcripts, ranging  in size from 200bp to 270bp (aveage = 250bp, 
  standard deviation = 17bp).  Each Illumina simulation produce 150K
  reads from the 100 zebra-fish transcript, each exactly 62bp.}\label{SRSim}
\end{table}

@@

% LocalWords:  overline TpG li lj lk ijk i'j'k jj kk trinucleotides rb Arndt et
% LocalWords:  transversion Arndt's dinucleotides transversions RepBase arallel
% LocalWords:  nalysis lustering ngine al substring Hazelhurst \textsc{wcd}  ESTsim ESTs
% LocalWords:  polymearse Methodoloy Jaccard tp fn fp  wcd th Prim's Ns ssembly
% LocalWords:  Chlamydomonas reinhartdii nvironment xpression runtime bp Runtme
% LocalWords:  stochastically Runtimes EasyCluster Ricinus Dataset

>>>>

\newcommand{\peace} {{\small PEACE}}
\newcommand{\wcd} {{\small WCD}}
\newcommand{\capthree} {{\small Cap3}}
\newcommand{\easycluster} {{\small EasyCluster}}
\newcommand{\metasim} {{\small MetaSim}}
\newcommand{\tgicl} {{\small TGICL}}

@@

In {\it Hazelhurst et al.} \cite{Hazelhurst08a} the authors conduct
an investigation of \textsc{wcd} \mc{(version 0.5.1)} against a number of clustering
tools.  As they make a convincing argument that \textsc{wcd} is returning
better results then the tools against which they compare, we limit our
analysis to a comparison of \textsc{peace} against \textsc{wcd},
\tgicl \cite{Pertea2003}, and the \textsc{cap3} assembly
tool \cite{Huang99} -- the last of which implicitly clusters while
performing assembly.

@@

For simulated data sets we use the {\bf ESTsim} tool to generate
simulated EST data sets \cite{Hazelhurst03}, using the collection of
zebra fish genes that served as a basis for the \textsc{wcd} simulations
\cite{Hazelhurst08a}.  In generating the ESTs, ESTsim models three
types of error (general base read errors, errors due to
polymerase decay, and primer interference), and allows those errors to
take the form of substitutions, deletions, and insertions (of bases
and Ns).  See the ESTSim paper for a discussion of the probability
distributions and default parameters.  For generating our simulated
data, we use the default values for all parameters that are not
explicitly being subjected to variation in our experiments --
parallelising the testing of the \textsc{wcd} tool.

@@

Each estimate given in the main paper, as well as the following, is
averaged over 30 trials.  Each trial consists of the application of
all tools to a simulated data set, the set having been derived
from the application of the ESTsim tool to a set of 100 zebra fish
gene sequences \cite{Hazelhurst03}.  All confidence intervals are
calculated at a 95\% level of significance.

@@

\mc{Our primary measurement tools are sensitivity, the Jaccard Index,
  Type 1 error and Type 2 error (see Figure~3 from the paper).
  Sensitivity is the ratio for true positives to actual positives ($tp
  / (tp + fn)$) -- the} fraction of sequences from the same transcript
that are identified as such.  The {\it Jaccard} index combines both
sensitivity and specificity, calculated as the ratio of true positives
to the sum of true positives, true negatives, and false positives ($tp
/ (tp+fn+fp)$) \cite{Hazelhurst08a}.

@@

\mc{Runtime of the tool is analyzed in Figure~4 of the main paper.  In
Figure~\ref{memory} we present a plot of the memory footprint of each tool
as a function of input size.  Data points were generated by the application of
each tool to simulated sets of a specified size (generated by ESTSim),
measuring the maximum amount of memory required for the run.  We see that while
all tools require memory linear in the input size, \wcd\/ requires
slightly less than does \peace, while both require significantly less
memory than \capthree.  In general we find the footprint of \peace\/
to be relatively small, as compared to other clustering tools
(discussed in {\it Hazelhurst et al.} \cite{Hazelhurst08a}).}

\mc{All runs were conducted on 3.0 GHz Intel Xeon EM64T CPUs with 2 MB
caches and a 800 MHz front side bus, model number E5520 (2005)}

\begin{figure}[tbp]
\centerline{
\includegraphics[scale=0.35]{pics.d/memory.pdf}
}
\caption{Plots of memory footprint against sequence set size for a
  single processor run on simulated data for each tool: \peace\/
  (blue), \wcd\/ (green) and \capthree\/ (black), along with the
  least-squares fit for each.  For each fit we have an associated
  regression coefficient of $r > 0.99$.}\label{memory}
\end{figure}

\subsection{Raising the Threshold}
\label{sim_thresh}
\mc{ 
  The \wcd\/ tool uses a $d^2$ threshold of $40$ to determine whether
  two sequences are in the same cluster.  As this is equivalent to our
  $T$ threshold for dismissing MST edges, we have set it to the same
  default value when used for Sanger} \mc{sequences.}  \mc{However,} \mc{we} \mc{did}
  \mc{experiment} \mc{with} \mc{higher} \mc{values,} \mc{with} \mc{interesting} \mc{results.  An
  analysis of the performance of $d^2$ indicated that a $T$ value as high as 130 should not have a
  significant impact on its effectiveness as a discriminator.  And in
  fact this was born out in simulation, as shown in
  Figure~\ref{SeJiT1T2130}.  In contrast with the equivalent results
  for a threshold of $T=40$, as shown in the manuscript, we see
  considerable improvement in \peace\/ over \wcd\/ in all metrics
  except Type 2 error, in which \wcd\/ slightly
  better. Unsurprisingly, this also makes a significant difference in
  the tool's ability to distinguish duplications: we see fairly poor
  performance in Figure~\ref{dups130}.  
}

@@

\begin{figure}[tbh]
  \centerline{\includegraphics[width=3.25in]{pics.d/SeJiT1T2x130.pdf}}
  \caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
    Type 1 error using the elevated threshold ($T=130$), providing a
    contrast against Figure~3 ($T=40$).  Based on the average over 30
    simulated Sanger Sequence ESTs sets derived from 100 zebra fish
    genes.  Blue/Solid = \peace, Green/Dash = \wcd, Black/Dot-Dash =
    \capthree; vertical tics = 95\% confidence intervals on estimates.
    Intervals are not presented for Type 1 error due to the effective
    lack of variance.}\label{SeJiT1T2130}
\end{figure}

\begin{figure}[tbh]
\centerline{
\includegraphics[scale=0.35]{pics.d/duplicates_130.pdf}
}
\caption{Ability to distinguish duplicates as a function of
  divergence.  Estimates averaged over 30 trials; one trial consists
  of taking a random zebra fish gene, copying it and stochastically changing bases
  at the specified rate, then using the two genes as the bases for
  generating a simulated set. Blue/Solid = \textsc{peace}, Green/Dashed = \textsc{wcd}, Black/Dot-Dashed = \textsc{cap3};
  variance was too small for visible plotting of
  confidence intervals.}\label{dups130}
\end{figure}

\subsection{Comparison against \tgicl}

In comparing against \wcd\/ we see equivalent results,
while comparisons against \capthree\/ show improvement in most of the
quality measures (at the cost of an ability to distinguish
duplicates).  Comparisons against \tgicl\/ \cite{Pertea2003} are less
straight forward.  In Figure~\ref{TGICL1} and Figure~\ref{TGICL2} we
repeat out plots from Figure 3 of the manuscript at each threshold
value.  We see that \peace\/ is more competitive with \tgicl\/ at higher
error rates in terms of sensitivity.  More generally, we see \peace\/
does a better job at specificity, as reflected both in the Jaccard index
(where the shift in the \tgicl\/ curve indicates a problem with type 2 errors) and
the higher Type 2 error rate.  We also see this reflected in the
ability of \tgicl\/ to distinguish duplicated genes, appearing to be unable
to consistently do so at a similarity of more than 72\% (as opposed to
about 80\% for \peace), shown in Figure~\ref{tgicl_dups}.

In a sequential execution, \tgicl\/ is clearly the faster tool --
though this does not necessarily not compensate for the lower
specificity and improved results of \peace\/ on short-read data.  While
both tools enable parallal computations, they are implemented on
different libraries: \peace\/ uses the standard OpenMPI library, while
\tgicl\/ is built on the PVM.  While MPI has been adapted as the
standard by IEEE and is generally availabel by default, PVM is not so
wide spread -- making \tgicl\/ less portable to new
systems in its parallel form.

\begin{figure}[tbh]
\centerline{
  \centerline{\includegraphics[width=3.35in]{pics.d/TGICL_SeJiT1T2x40.png}}
}
\caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
  Type 1 error, based on the average over 30 simulated Sanger Sequence
  ESTs sets derived from 100 zebra fish genes, using a \peace\/ threshold
  of $T=40$.  Blue/Solid = \peace, Green/Dash = \wcd\/ \mc{(version
    0.5.1)}, Black/Dot-Dash = \capthree; Purple/Long-Dash = TGICL;
  vertical tics = 95\% confidence intervals on estimates.  Intervals
  are not presented for Type 1 error due to the effective lack of
  variance.}
\label{TGICL1}
\end{figure}

\begin{figure}[tbh]
\centerline{
  \centerline{\includegraphics[width=3.35in]{pics.d/TGICL_SeJiT1T2x130.png}}
}
\caption{Comparisons of Sensitivity, Jaccard Index, Type 1 error and
  Type 1 error, based on the average over 30 simulated Sanger Sequence
  ESTs sets derived from 100 zebra fish genes, using a \peace\/ threshold
  of $T=130$.  Blue/Solid = \peace, Green/Dash = \wcd\/ \mc{(version
    0.5.1)}, Black/Dot-Dash = \capthree; Purple/Long-Dash = TGICL;
  vertical tics = 95\% confidence intervals on estimates.  Intervals
  are not presented for Type 1 error due to the effective lack of
  variance.}
\label{TGICL2}
\end{figure}

\begin{figure}[tbh]
\centerline{
  \centerline{\includegraphics[width=3.35in]{pics.d/tgicl_duplicates_40.png}}
}
\caption{Ability to distinguish duplicates as a function of divergence
  (\peace\/ v. \mbox{\tgicl}).  Estimates averaged over 30 trials; one trial
  consists of taking a random zebra fish gene, copying it and
  stochastically changing bases at the specified rate, then using the
  two genes as the bases for generating a simulated set. Blue/Solid =
  \peace, Purple/LongDash = \tgicl; variance was too small for visible
  plotting of confidence intervals.}\label{tgicl_dups}
\end{figure}

\section{Benchmark Data}

@@

\begin{table}[tbh]
\begin{center}
\begin{tabular} {|c|| c | c || c | c |}
\hline
 Benchmark & \begin{tabular}{c} Number of \\ Sequences \end{tabular}
 & \begin{tabular}{c} Number of \\ Bases \end{tabular} & Peace Runtme (s)&
 WCD Runtime (s) \\
\hline
\hline
 Human Cluster & 17,733 & 11.7 Mb & 293 & 804 \\
\hline
 Cotton & 29,995 & 17.1 Mb & 222 & 339 \\
\hline
 Ricinus & 57,690 & 40.9 Mb & 1344 & 2356 \\
\hline
 A076941 & 76,941 & 32.8 Mb & 1166 & 966 \\
\hline
\hline
 A686903 & 686,904 & 294.7 Mb & 9449 & 2481 \\
\hline
\end{tabular}
\end{center}
\caption{Runtime for \peace\/ and \wcd\/ on benchmarks from the
  WCD and EasyCluster studies \cite{Hazelhurst08a, Picardi09}.
  Runtimes for the first four sets are sequential, while that last was
  run over thirty processors.}\label{RunTime}
\end{table}

@@

\begin{table}[tbh]
\begin{center}
\begin{tiny}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime
      (s) \end{tabular} \\
\hline
\hline EasyCluster Human & PEACE & 0.999 & 0.106 & 0.027 & 0.030 & 67 & 2 & 295 \\
\cline{2-9} Benchmark & WCD & 0.998 & 0.672 & 0.144 & 0.044 & 113 & 16 & 804 \\
\cline{2-9} (111 Genes) & Cap3 & 0.657 & 0.643 & 1.000 & 0.001 & 2269 & 1827 & \mbox{NA} \\
\hline
\hline WCD A076941 & PEACE & 0.937 & 0.345 & 0.337 & 0.0342 & 18256 & 8135 & 1293 \\
\cline{2-9} Benchmark & WCD & 0.933 & 0.476 & 0.350 &  0.027 & 18787 & 8553 & 966 \\
\cline{2-9} (13240 genes) & Cap3 & 0.826 & 0.802 & 0.486 & 0.014 & 25042 & 14916 &
\mbox{NA} \\
\hline
\end{tabular}
\end{tiny}
\end{center}
\caption{Comparisons of runs on the EasyCluster human Benchmark
  Dataset and the WCD A076941 Arabidopsis thaliana dataset using the
  standard quality measurements with $T=130$.}\label{SangerReal130}

@@

\begin{table}[htbp]
\begin{center}
\begin{tiny}
\begin{tabular} {| c || c || c | c | c | c | c | c | c |}
\hline
 & 
& Sensitivity & Jaccard & \begin{tabular}{c} Type 1 \\
  error \end{tabular} & \begin{tabular}{c} Type 2 \\
  error \end{tabular} &
\begin{tabular}{c} Number \\ of  \\
    Clusters \end{tabular} & \begin{tabular}{c} Number \\ of \\
    Singletons \end{tabular}
    & \begin{tabular}{c} Single \\ processor \\ runtime (s) \end{tabular}\\
\hline
\hline   454 & PEACE & 0.871 & 0.809 & 0.186 & 0.049 & 129 & 4 & 129 \\
\cline{2-9}    Simulation & WCD(1) & 0.459 & 0.448 & 0.946 & 0.003 & 599 & 235 & 76 \\
\cline{2-9}  & TGICL & 0.297 & 0.288 & 0.984 & 0.005 & 886 & 0 & 68 \\
\hline
\hline Illumina         & PEACE & 0.384 & 0.343 & 0.230 & 0.018 & 673  & 299 & 3463 \\
\cline{2-9} Simulation  & TGICL & 0.131 & 0.127 & 0.977 & 0.008 & 1522 & 0   & 373 \\
\hline
\end{tabular}
\end{tiny}
\end{center}
\caption{Comparisons of runs on on short read data, as generated by
  \metasim \cite{Richter2008}; each figure is an average over 30 runs.
  For \wcd we use parameters suggested by the authors  (``-H 1 -T 160''). \wcd\/ was
  unable to return any reasonable results for Illumina data, while
  \capthree\/ was unable to return results for either simulation.  Each
  454 simulation produced 25,000 reads from the 100 zebra-fish
  transcripts, ranging  in size from 200bp to 270bp (average = 250bp, 
  standard deviation = 17bp).  Each Illumina simulation produce 150K
  reads from the 100 zebra-fish transcript, each exactly 62bp.}\label{SRSim}
\end{table}

@@

% LocalWords:  overline TpG li lj lk ijk i'j'k jj kk trinucleotides rb Arndt et
% LocalWords:  transversion Arndt's dinucleotides transversions RepBase arallel
% LocalWords:  nalysis lustering ngine al substring Hazelhurst \textsc{wcd}  ESTsim ESTs
% LocalWords:  polymearse Methodoloy Jaccard tp fn fp  wcd th Prim's Ns ssembly
% LocalWords:  Chlamydomonas reinhartdii nvironment xpression runtime bp Runtme
% LocalWords:  stochastically Runtimes EasyCluster Ricinus Dataset  GHz Xeon LV
% LocalWords:  CPUs TGICL Arabidopsis thaliana dataset Illumina ESTSim LongDash
% LocalWords:  parallelising
